---
title: "AI Activities"
listing:
  contents: ai
  sort: title
  sort-ui: false
  filter-ui: false
  type: table
  fields: [title, subtitle, description]
  field-display-names:
    title: "No."
    subtitle: "Title"
    description: "Content"
---

-   Activity 1 is presented by [**Superb Statisticians**]{.blue}: Aaron, Connor, and Danielle.

-   Activity 2 is presented by [**Team 1**]{.blue}: Sophia, Maya, Kaitlyn.

```{r}
teams <- c("Team 1", "The Data Scientists", "Stat Padders", "Data Divers", "")
set.seed(3570)
sample(teams, 1)
```

-   Activity 3 is presented by team [**The Data Scientists**]{.blue}: Eliza, Miranda, Lauren.

```{r}
teams <- c("The Data Scientists", "Stat Padders", "Data Divers", "Plot Squad")
set.seed(3570)
sample(teams, 4)
```

-   Activity 4 is presented by team [**Stat Padders**]{.blue}:  Jared, DJ, Raj.

-   Activity 5 is presented by team [**Plot Squad**]{.blue}: Dominic, Wyatt, Drishtant, Tyler.

-   Activity 6 is presented by team [**Data Divers**]{.blue}: Karlo, Beth, Jack.

