---
title: "Lab 10: Import Data"
editor: source
reference-location: margin
execute:
  freeze: true
  echo: true
---

```{r}
#| echo: false
knitr::opts_chunk$set(
    fig.asp = 0.618,
    fig.align = "center",
    out.width = "100%",
    fig.retina = 10,
    fig.path = "../slides/images/08-import/",
    message = FALSE,
    global.par = TRUE
)
library(fontawesome)
```

::: callout-note

:::


- If you haven't, install and load the `tidyverse` package.

In **lab.qmd** `## Lab 10` section,
<!-- - If you haven't, upload data.zip posted in D2L Week 5 module to your lab project workspace. -->
<!-- - Run the following code to see the data sets saved in the `dslabs` package: -->
<!-- ```{r, eval=FALSE} -->
<!-- path <- system.file("extdata", package = "dslabs") -->
<!-- files <- list.files(path) -->
<!-- files -->
<!-- ``` -->
<!-- - Use `file.path()` to create the file path for the file **"ssa-death-probability.csv"**.  -->


- Import **ssa_male_prob.csv** and **ssa_female_prob.Rds** in the data folder using `read_csv()` and call them `ssa_male` and `ssa_female`, respectively.

```{r}
#| eval: false
#| purl: false
ssa_male <- readr::read____(____________)
ssa_female <- readr::read____(____________)
```

- Plot `Age` (x-axis) vs. `LifeExp` (y-axis) for `Female`. The type should be "line", and the line color is red. Add x-label, y-label and title to your plot.

```{r}
#| eval: false
#| purl: false
plot(x = _____, y = _____, type = ______, col = ______,
     xlab = ______, ylab = _______, main = ____________)
```

- Use `lines()` to add a line of `Age` (x-axis) vs. `LifeExp` (y-axis) for `Male` to the plot. The color is blue.

```{r}
#| eval: false
#| purl: false
lines(x = _____, y = _____, col = ______)
```