---
title: "Mini Project 2"
subtitle: "Data Wrangling, Reproducibility, and Pipelines"
format: html
editor: source
categories: "Mini Project"
date-modified: today
execute:
  echo: true
  eval: true
  warning: false
  message: false
---

```{r}
#| echo: false
library(tidyverse)
```

## Overview

In **Mini Project 1**, our team explored a data set and told a short story with visuals. In this second mini project, we focus on a different but equally important part of data science: **turning messy data into a clean, analysis ready table**.

For this project, we use the built in `billboard` data set from the `tidyr` package. The main wrangling challenge is that the weekly chart rankings are stored across many columns (`wk1`, `wk2`, `wk3`, and so on), which makes the data hard to analyze directly. Our goal is to build a clear pipeline that reshapes and cleans the data so that one row represents **one song in one week on the chart**.

This is a **data wrangling project**. We focus on **cleaning, organizing, and documenting** the data. We keep the final analysis brief and descriptive.

### What we will submit

In Posit Cloud, we will show

1. The **rendered HTML report**
2. The **complete version of this source qmd file**
3. The raw data source, which in this case is the built in `billboard` data set from `tidyr`
4. One exported cleaned data file named **`data_clean.csv`**

### What we will present

A **10 minute** team presentation that explains

1. What was messy or unclear in the original data
2. What our cleaning pipeline did
3. Why the final cleaned table is more trustworthy and useful

## Team Info

- **Team Name**: Data Divers

- Team members and roles for this project:

  1. **Project lead** (keeps time, coordinates tasks): Alex
  2. **Pipeline builder** (writes and organizes the cleaning code): Brianna
  3. **Quality checker** (checks missing values, duplicates, odd values): Carlos
  4. **Documentation writer** (explains steps and prepares presentation): Dana

Each student has a specific role for accountability, but we worked together throughout the project.

## Step 1: Choose a data set

We chose the `billboard` data set from the `tidyr` package.

```{r}
#| label: import-data
data_raw <- tidyr::billboard
```

### Quick description of your data set

1. What does one row represent

One row in the raw data represents one song that appeared on the Billboard chart.

2. What makes this data set messy, incomplete, or hard to use right away

The weekly chart rankings are spread across many columns such as `wk1`, `wk2`, `wk3`, and so on. This wide format makes it difficult to answer questions about a song's position over time. Also, there are many `NA` values in later week columns because songs often leave the chart before week 76.

3. What might someone want to do with the cleaned version of this data

Someone might want to study how long songs stay on the chart, how rankings change over time, or which songs had the strongest chart performance.

4. If you are using more than one raw table, what does each table contribute

We are using only one raw table for this project.

## Step 2: Mini proposal

1. What is the goal of your cleaning pipeline

The goal is to reshape the data so that each row represents one song in one chart week, with a clear week number, chart rank, and date.

2. Who would benefit from having this cleaned table

A data analyst, a music researcher, or a student learning about tidy data would benefit from having a cleaned version of this table.

3. List at least 4 wrangling tasks you expect to do

- Rename unclear variables
- Convert the chart entry date to a date format
- Pivot the weekly ranking columns into a long format
- Extract numeric week values from the week column names
- Remove rows where the song was not on the chart that week
- Create a new date variable for each week on the chart

4. What will your final cleaned table allow someone to do more easily

The cleaned table will make it easier to track a song across weeks, summarize chart performance, and create plots or tables based on time on the chart.

5. What is one challenge you expect to face

One challenge is deciding how to interpret the many `NA` values in the weekly ranking columns. In this case, those `NA` values are meaningful because they show that a song was no longer on the chart.

## Step 3: Quick raw data check

Before cleaning, inspect the data and identify problems.

### 3.1 First look

```{r}
#| label: raw-check
glimpse(data_raw)
dim(data_raw)
names(data_raw)[1:15]
```

In the raw data, one row represents one song, but the weekly ranks are spread across many columns. This means the structure is wide rather than tidy. The data are usable for storage, but not convenient for analysis of change over time.

### 3.2 Missing values

```{r}
#| label: missing-values
missing_counts <- data_raw |>
  summarise(across(everything(), ~ sum(is.na(.x)))) |>
  pivot_longer(cols = everything(), names_to = "variable", values_to = "n_missing") |>
  arrange(desc(n_missing))

missing_counts
```

Most of the missing values appear in later weekly chart columns. These missing values do not appear to be errors in data entry. Instead, they mean that the song was no longer on the chart during those later weeks. Because our final cleaned table should represent actual chart appearances, we plan to remove rows with missing ranks after we reshape the data.

### 3.3 Duplicate rows or repeated records

```{r}
#| label: duplicates
nrow(data_raw)

nrow(distinct(data_raw))

data_raw |>
  count(artist, track, sort = TRUE) |>
  filter(n > 1)
```

The number of rows in `data_raw` is the same as the number of distinct rows, so there are no exact duplicate rows. Each song appears once in the raw table.

### 3.4 Other issues

```{r}
#| label: other-issues
names(data_raw) |>
  tail(10)
```

The weekly ranking columns are named `wk1`, `wk2`, and so on, which means the week number is stored inside the column name rather than in its own variable. This is a classic tidy data problem because one variable, week, is spread across many columns.

## Step 4: Build your cleaning pipeline

Create a new object `data` that is ready for later analysis.

Our pipeline includes more than 4 meaningful wrangling actions:
1. Renaming variables
2. Converting a text date to a date object
3. Pivoting the weekly columns from wide to long
4. Extracting numeric week values
5. Removing rows with missing ranks
6. Creating a date for each week on the chart
7. Arranging rows in a clean order

```{r}
#| label: clean
data <- data_raw |>
  rename(
    artist_name = artist,
    song_title = track,
    chart_entry_date = date.entered
  ) |>
  mutate(
    chart_entry_date = as.Date(chart_entry_date, format = "%m/%d/%Y")
  ) |>
  pivot_longer(
    cols = starts_with("wk"),
    names_to = "week",
    values_to = "rank"
  ) |>
  mutate(
    week = readr::parse_number(week),
    rank = as.integer(rank),
    date_on_chart = chart_entry_date + 7 * (week - 1)
  ) |>
  filter(!is.na(rank)) |>
  arrange(artist_name, song_title, week)

data
```

### Describe your main cleaning steps

1. **Step 1:** We renamed variables such as `artist`, `track`, and `date.entered` so that the names are more descriptive and easier to read.

2. **Step 2:** We converted `chart_entry_date` from text into a true date object so that R can recognize and work with it as a calendar date.

3. **Step 3:** We used `pivot_longer()` to turn the many `wk` columns into two variables: `week` and `rank`.

4. **Step 4:** We extracted the numeric week number from values like `wk1` and `wk2`, then created a new variable called `date_on_chart` to show the date associated with each chart week.

5. **Optional additional step:** We removed rows with missing ranks because those rows do not represent actual chart appearances.

### Why your pipeline is readable

We used the pipe operator so that each step follows logically from the previous one. We also grouped related tasks together, such as renaming, reshaping, and date creation. This makes the code easier for another person to follow from top to bottom.

## Step 5: Check your cleaned data

Now verify that your cleaned table makes sense.

### 5.1 Final structure

```{r}
#| label: final-structure
glimpse(data)
summary(data)
```

The cleaned table is much easier to understand because each row now represents one song in one week on the chart. The `week` variable is explicit, the `rank` variable is stored in one column, and the date information is easier to interpret.

### 5.2 Quality checks

We run 4 simple checks:
1. Missing values in `rank`
2. Duplicate song and week combinations
3. Range of chart ranks
4. Range of week values

```{r}
#| label: quality-checks
check_missing_rank <- sum(is.na(data$rank))

check_duplicate_song_week <- data |>
  count(artist_name, song_title, week) |>
  filter(n > 1) |>
  nrow()

check_rank_range <- range(data$rank)

check_week_range <- range(data$week)

quality_results <- tibble(
  check = c(
    "Missing values in rank",
    "Duplicate song-week rows",
    "Minimum and maximum rank",
    "Minimum and maximum week"
  ),
  result = c(
    check_missing_rank,
    check_duplicate_song_week,
    paste(check_rank_range, collapse = " to "),
    paste(check_week_range, collapse = " to ")
  )
)

quality_results
```

These checks suggest that the cleaned data are reasonable. There are no missing values in `rank` after filtering, there are no duplicate song and week combinations, and the rank values and week values fall in sensible ranges.

### 5.3 Export your cleaned data

```{r}
#| label: export-cleaned
write_csv(data, "data_clean.csv")
```

## Step 6: Show that your table is analysis ready

### 6.1 Preview of cleaned data

```{r}
#| label: preview-cleaned
slice_head(data, n = 10)
```

This table is more usable than the raw version because the weekly chart information is now stored row by row rather than across many columns. That makes it much easier to summarize, filter, and visualize a song's performance over time.

### 6.2 Mini data dictionary

```{r}
#| label: data-dictionary
data_dictionary <- tibble(
  variable = c(
    "artist_name",
    "song_title",
    "chart_entry_date",
    "week",
    "rank",
    "date_on_chart"
  ),
  meaning = c(
    "Name of the artist",
    "Title of the song",
    "Date the song first entered the chart",
    "Week number since the song entered the chart",
    "Chart rank in that week",
    "Calendar date corresponding to that chart week"
  ),
  type = c(
    "character",
    "character",
    "date",
    "integer",
    "integer",
    "date"
  )
)

data_dictionary
```

### 6.3 One simple summary that the cleaned table now supports

We now create a summary table of the songs that stayed on the chart the longest.

```{r}
#| label: simple-summary
song_summary <- data |>
  group_by(artist_name, song_title) |>
  summarise(
    weeks_on_chart = n(),
    best_rank = min(rank),
    .groups = "drop"
  ) |>
  arrange(desc(weeks_on_chart), best_rank)

slice_head(song_summary, n = 10)
```

This summary shows which songs remained on the chart for the most weeks and what their best chart rank was. This would have been much harder to calculate from the raw wide format because the weekly information was spread across many columns.

### 6.4 Handoff note

This cleaned table is designed for someone who wants to analyze chart performance over time. Each row represents one song in one week on the Billboard chart. The most important variables are `artist_name`, `song_title`, `week`, `rank`, and `date_on_chart`. One limitation is that this table only includes songs that made the chart and only tracks ranks while they remained on the chart. A reasonable next step would be to study how quickly songs rise or fall in ranking over time.

## Step 7: Pipeline quality checklist

We confirm that our report shows the following:

1. We clearly explained what was messy in the raw data
2. Our cleaning code can be rerun from top to bottom
3. Our cleaned table is saved as `data`
4. We exported `data_clean.csv`
5. We ran at least 3 checks on the final data
6. Our explanations are in plain language
7. We did not overcomplicate the project

## Step 8: Team reflection

Each team member writes 2 to 4 sentences.

#### Member 1: Alex

I helped coordinate the project and kept track of the required sections. I learned that good data science starts before analysis, because the structure of the data can either make the work easy or very confusing. Next time, I would spend a little more time planning the quality checks before writing code.

#### Member 2: Brianna

I built most of the wrangling pipeline. I learned how useful `pivot_longer()` is when the same variable is spread across many columns. Next time, I would add more comments while coding so the logic is even clearer.

#### Member 3: Carlos

I focused on checking missing values, duplicates, and value ranges. I learned that not all missing values mean the same thing, and that sometimes `NA` has a meaningful interpretation. Next time, I would create a more formal checklist before doing the final review.

#### Member 4: Dana

I wrote the explanations and organized the report. I learned that documentation matters because someone else needs to understand the cleaned data later. Next time, I would make the data dictionary earlier in the project instead of waiting until the end.

## Step 9: Presentation plan

We plan a **10 minute** talk with this structure:

1. About 1 minute: context and purpose of the cleaned table
2. About 2 minutes: what was messy or unclear in the raw data
3. About 3 minutes: our main wrangling steps
4. About 2 minutes: our quality checks and what they showed
5. About 2 minutes: the final cleaned table, next use, and takeaway

### Presentation order

```{r}
#| eval: false
teams <- c(
  "Team 1",
  "Superb Statisticians",
  "The Data Scientists",
  "Stat Padders",
  "Data Divers",
  "Plot Squad"
)

set.seed(3570)
sample(teams, 6, replace = FALSE)
```

## Grading guide

Total 15 points:

1. **Clear** purpose and clear identification of raw data issues (3 pts)
2. **Meaningful** wrangling pipeline and clear documentation of steps (6 pts)
3. **Reproducible** workflow, quality checks, and usable final cleaned table (3 pts)
4. Presentation **clarity** and **timing** (3 pts)