---
title: "Mini Project 3"
subtitle: "Supervised Learning and Model Evaluation"
format: html
editor: source
categories: "Mini Project"
date-modified: today
execute:
  echo: true
  eval: false
  warning: false
  message: false
---

```{r}
#| echo: false
library(tidyverse)
```

## Overview

In **Mini Project 1**, your team explored data and told a short story with visuals. In **Mini Project 2**, your team built a cleaning pipeline and created a trustworthy, analysis ready table.

In **Mini Project 3**, your team will take one more step: **use data to make a prediction and evaluate how well it works on new data**.

Complete the Quarto file by

1. Defining a clear prediction task
2. Choosing a target variable and a small set of predictors
3. Splitting the data into a training set and a test set
4. Fitting **2 simple supervised learning models**
5. Comparing model performance on the test set
6. Explaining results and limitations in plain language

Keep the project simple and honest. Do NOT do advanced tuning, deep learning, or a large number of models.

## What you will submit

Show, in your Posit Cloud,

1. The **rendered HTML report**
2. The **complete version of this source qmd file**
3. **Any data file you used**, only if it is not a built in package data set
4. Any cleaned data file you created for this project, if applicable

## What you will present

A **10 minute** team presentation that explains

1. What your prediction task was
2. Which variables you used
3. How you split the data
4. Which 2 models you compared
5. Which model performed better on the test set
6. What your team learned and what the model cannot tell us

## Team Info

- **Team Name**: [Your Team Name]{.blue}

- Team members and roles for this project:

  1. **Project lead** (keeps time, coordinates tasks): [Your Member Name(s)]{.blue}
  2. **Data and feature lead** (prepares target and predictors): [Your Member Name(s)]{.blue}
  3. **Modeling lead** (fits models and organizes output): [Your Member Name(s)]{.blue}
  4. **Evaluation and presentation lead** (compares models and prepares presentation): [Your Member Name(s)]{.blue}

[Role note]: Each student plays a specific role for accountability, but teammates are welcome to work together, help, and learn from each other to produce high quality and cohesive work. If your team has 3 members, one person may take on two roles.

## Project rules

Keep the project simple.

1. Choose **one** prediction question
2. Use **one** data set
3. Use **no more than 6 predictors**
4. If you do a classification project, choose a **binary** target variable
5. Fit **exactly 2 simple models**
6. Use **one** training and test split
7. Use **1 or 2 evaluation metrics**
8. Explain results in plain language
9. Do NOT claim causation
10. Do NOT overstate what your model can do

## Step 1: Choose a data set

Choose one option below. Two teams may use the same data set.

### Data sets

1. `penguins` data set from R package [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/)
2. `mpg` data set from R package [`ggplot2`](https://ggplot2.tidyverse.org/reference/mpg.html)
3. `loans_full_schema` data set from R package [`openintro`](https://www.openintro.org/)
4. Your cleaned data from **Mini Project 2**, with instructor approval
5. Your own data set, with instructor approval

:::: callout-warning
::: {style="font-size: 1.1em;"}

Choose a data set and target that are manageable for a short team project. Your goal is not to build the most accurate model possible. Your goal is to demonstrate a clear and correct prediction workflow.

:::
::::

Use the code chunk `import-data` below to import your data, and call the main table `data_raw`.

```{r}
#| label: import-data
## Import your data here
## Example:
## data_raw <- palmerpenguins::penguins
data_raw <- 
```

### Quick description of your data set

Answer the following questions.

1. What does one row represent

[**Answer:**]{.blue}

2. Why is this data set suitable for a prediction task

[**Answer:**]{.blue}

3. What variable do you want to predict

[**Answer:**]{.blue}

4. Is your task a **regression** problem or a **classification** problem

[**Answer:**]{.blue}

## Step 2: Mini proposal

Write short answers. Keep them specific.

1. What is your prediction question

[**Answer:**]{.blue}

2. Who might care about this prediction question

[**Answer:**]{.blue}

3. What is your target variable

[**Answer:**]{.blue}

4. Which predictors do you plan to use, and why

[**Answer:**]{.blue}

5. Why is this a reasonable introductory project

[**Answer:**]{.blue}

6. What is one challenge or limitation you expect

[**Answer:**]{.blue}

## Step 3: Prepare your modeling data

Create a table called `data` that is ready for modeling.

Keep this preparation light and focused. You may

1. Select useful variables
2. Filter rows
3. Handle missing values
4. Recode values
5. Create a small number of simple derived variables

Do NOT turn this into another full wrangling project. If major cleaning was already done in **Mini Project 2**, briefly summarize that and then move on.

```{r}
#| label: prepare-data
## Prepare your modeling data here
data <- 
```

### Describe your modeling data

1. How many rows are in `data`

[**Answer:**]{.blue}

2. What is the target variable

[**Answer:**]{.blue}

3. Which predictors did you keep

[**Answer:**]{.blue}

4. Did you remove any rows or variables, and why

[**Answer:**]{.blue}

## Step 4: Quick check of the target and predictors

Before fitting models, inspect your variables.

```{r}
#| label: quick-check
## Check your target and predictors here
## Suggestions:
## glimpse(data)
## summary(data)
```

### Target check

Describe the target variable.

1. If regression, what is its range and general distribution
2. If classification, what are the class counts

[**Answer:**]{.blue}

### Predictor check

Describe any issues you noticed with the predictors, such as missing values, unusual values, or highly unbalanced categories.

[**Answer:**]{.blue}

## Step 5: Split the data into training and test sets

Use one train and test split. Set a seed so that the split can be reproduced.

A common choice is about **80 percent** for training and the rest for testing.

```{r}
#| label: split-data
set.seed(3570)

## Create training and test sets here
## Example logic:
## n <- nrow(data)
## train_id <- sample(seq_len(n), size = round(0.75 * n))
## train <- data[train_id, ]
## test  <- data[-train_id, ]
```

### Why do we split the data

In 2 to 4 sentences, explain why the test set is important.

[**Answer:**]{.blue}

### Split summary

1. Number of rows in training set: [**Answer:**]{.blue}
2. Number of rows in test set: [**Answer:**]{.blue}

If classification, report the class counts in both sets.

[**Answer:**]{.blue}

## Step 6: Model 1

Choose a simple first model.

### Recommended Model 1

1. If regression, use **linear regression**
2. If classification, use **logistic regression**

If you want to use a different Model 1, get instructor approval.

### Model 1 formula

Write your model formula in words.

[**Answer:**]{.blue}

```{r}
#| label: fit-model-1
## Fit Model 1 here
model_1 <- 
```

### Model 1 interpretation

Briefly describe what Model 1 is doing.

[**Answer:**]{.blue}

## Step 7: Model 2

Choose a second simple model that is different from Model 1.

### Recommended Model 2

1. If regression, use a **decision tree for regression**
2. If classification, use a **decision tree for classification**

If you want to use a different Model 2, get instructor approval.

### Why choose this second model

[**Answer:**]{.blue}

```{r}
#| label: fit-model-2
## Fit Model 2 here
model_2 <- 
```

### Model 2 interpretation

Briefly describe what Model 2 is doing.

[**Answer:**]{.blue}

## Step 8: Make predictions on the test set

Use both models to predict the target variable on the test set.

```{r}
#| label: predictions
## Create predictions from both models here
## Suggested output names:
## pred_1
## pred_2
```

## Step 9: Evaluate model performance

Use the test set only.

### If your project is regression

Choose **1 or 2** of the following:

1. RMSE (root mean square error)
2. MAE (mean absolute error)
3. $R^2$

Also include **one simple plot** such as predicted versus actual values.

### If your project is classification

Choose **1 or 2** of the following:

1. Accuracy
2. Misclassification rate
3. Sensitivity
4. Specificity

Also include a **confusion matrix**.

```{r}
#| label: evaluate-models
## Evaluate Model 1 and Model 2 here
## Report your test set performance clearly
```

### Results summary

Fill in the performance of both models.

#### Model 1

[**Answer:**]{.blue}

#### Model 2

[**Answer:**]{.blue}

### Which model did better

[**Answer:**]{.blue}

### Was the difference large or small

[**Answer:**]{.blue}

## Step 10: Show one helpful output

Create **one** output that helps the audience understand the model results.

Examples:

1. A confusion matrix
2. A predicted versus actual plot
3. A small table comparing metrics
4. A simple visualization of prediction errors

```{r}
#| label: helpful-output
## Add one helpful output here
```

Explain why this output helps the audience understand the results.

[**Answer:**]{.blue}

## Step 11: Interpret the results in plain language

Answer the following questions.

1. What did your team learn from this prediction task

[**Answer:**]{.blue}

2. What can your model do reasonably well

[**Answer:**]{.blue}

3. Where might your model fail or be less reliable

[**Answer:**]{.blue}

4. What should we be careful **not** to claim from this project

[**Answer:**]{.blue}

5. If you had more time, what is one reasonable next step

[**Answer:**]{.blue}

## Step 12: Model honesty checklist

Confirm that your report shows the following:

1. You clearly defined a prediction question
2. You identified the target and predictors
3. You used a reproducible train and test split
4. You fit **exactly 2** models
5. You evaluated performance on the **test set**
6. You explained results in plain language
7. You discussed at least one limitation
8. You did not overclaim what the model means

Revise your report if needed.

## Step 13: Team reflection

Each team member writes 2 to 4 sentences:

1. What you contributed
2. One thing you learned about supervised learning
3. One thing you would improve next time

#### Member 1: your name

[**Answer:**]{.blue}

#### Member 2: your name

[**Answer:**]{.blue}

#### Member 3: your name

[**Answer:**]{.blue}

#### Member 4: your name (if applicable)

[**Answer:**]{.blue}

## Step 14: Presentation plan

Plan a **10 minute** talk with the suggested structure:

1. About 1 minute: data set and prediction question
2. About 2 minutes: target, predictors, and data preparation
3. About 2 minutes: train and test split
4. About 2 minutes: Model 1 and Model 2
5. About 2 minutes: test set results and comparison
6. About 1 minute: takeaway and limitations

### Presentation order

```{r}
#| eval: true
teams <- c("Team 1","Superb Statisticians", "The Data Scientists", "Stat Padders", "Data Divers", "Plot Squad")
set.seed(19)
sample(teams, 6, replace = FALSE)
```

## Grading guide

Total 15 points:

1. **Clear** prediction question, target, and predictors (3 pts)
2. **Reasonable** train and test workflow and appropriate model setup (4 pts)
3. **Correct** evaluation and honest comparison of the 2 models (4 pts)
4. **Clear** interpretation, limitations, and communication (4 pts)